Performance Evaluation of Big Data Management Strategies\\ for Neuroinformatics

### 2018-11-21 paper.tex
A few more comments from @ValHayot's course project report
Typo
Wrote the page cache model

% Our question: is it relevant to use Big Data engines for neuroinformatics? 
% Implications: migrate existing pipelines, adopt a new framework (costly)
% Some papers said yes, for instance Ariel's and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
% Perhaps this too: https://www.sciencedirect.com/science/article/pii/S0925231218311470 Spark is being used to implement neuroinformatics pipelines, but is it
% the right tool?
% But a quantitative evaluation is lacking.

@@

@@

%% Main concepts
% detailed description relevant to the experiment
% how shuffle works
% how data locality is implemented
% spills to disk when dataset (and derived data) is too large for memory
% Lazy evaluation
% overlay cluster: not considered here as this would be scheduling

@@

% general description relevant to the experiment
% two plugins
% provenance
% memory management (bad?)

@@

\subsection{Models: Write Buffer} % perhaps other models if relevant. 0.5 page

@@

@@

@@

% Future work: scheduling in a shared environment.

@@

\bibliographystyle{IEEEtran} 
\bibliography{IEEEfull,biblio}

>>>>

% Our question: is it relevant to use Big Data engines for neuroinformatics? 
% Implications: migrate existing pipelines, adopt a new framework (costly)
% Some papers said yes, for instance Ariel's, Thunder and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
% Perhaps this too: https://www.sciencedirect.com/science/article/pii/S0925231218311470 Spark is being used to implement neuroinformatics pipelines, but is it
% the right tool?
% But a quantitative evaluation is lacking.

@@

% define pipeline and workflows

@@

%% Main concepts
% pipeline description
% detailed description relevant to the experiment
% how shuffle works
% how data locality is implemented
% spills to disk when dataset (and derived data) is too large for memory
% Lazy evaluation
% persistence (refer to Thunder)
% multiple schedulers: we focus on Spark Standalone
% overlay cluster: not considered here as this would be scheduling

@@

% pipeline description: map nodes and their limitations
% general description relevant to the experiment
% two plugins
% provenance
% file system caching
% memory management (bad?)

@@

\subsection{Page Cache Model} % perhaps other models if relevant. 0.5 page

@@

The Linux kernel uses in-memory page caching to speed-up disk writes. 
When a process calls the \texttt{write()} system call, data is copied 
to a memory cache that is asynchronously written to disk by flusher 
threads when memory shrinks, ``dirty" (unwritten) data grows, or a 
process invokes the \texttt{sync()} system call~\cite{love2010linux}. 
The flushing of page cache to disk is also called the writeback 
process.

Page caching has a dramatic effect on I/O, reducing the duration of 
disk writes by several orders of magnitude. It is essentially a way to 
emulate in-memory computing at the kernel level, without requiring a 
dedicated engine. The size of the page cache, however, becomes a 
limitation when processes write faster than the disk bandwidth. When 
this happens, the page cache rapidly fills up and writes are limited by 
the disk write bandwidth -- typically 500~MB/s for random writes to 
contemporary SSDs  -- as if no page cache was involved.

We introduce the following basic model to describe the filling and 
flushing of the page cache by an application:
$$
d(t) = \left( \frac{D}{C} - \frac{\delta}{\gamma} \right)t + d_0,
$$
where:
\begin{itemize}
\item $d(t)$ is the amount of data in the page cache at time t
\item $D$ is the total amount of data written by the application
\item $C$ is the total CPU time of the application
\item $\delta$ is the disk bandwidth
\item $\gamma$ is the max number of concurrent processes
\item $d_0$ is the amount of data in the page cache at time $t_0$
\end{itemize}

This model applies to parallel applications assuming that (1) 
concurrent processes all write the same amount of data, (2) 
concurrent processes all consume the same CPU time, (3) data is written 
uniformly along task execution. Due these assumptions, all the 
processes will write at the same rate, which explains why the model 
does not depend on the total number of concurrent processes in the 
application, but only on the max number of concurrent processes 
executing on the same node ($\gamma$). We realize that these 
assumptions would usually be violated in practice, but this simple 
model already provides interesting insights on the performance of disk 
writes, as shown later. Naturally, the model also ignores other 
processes that might be writing to disk concurrently to the 
application, which we assume negligible here.

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-compute ratio beyond which the page cache becomes 
asymptotically useless, which we will use to define our benchmarks.

% Cache eviction: LRU/n.

@@

% Do we see the effect of Java serialization?

@@

% Future work:
% - scheduling in a shared environment.
% - workflow-aware cache eviction strategies instead of LRU/n

@@

\bibliographystyle{IEEEtran} 
\bibliography{biblio}
